---
title: 'Statistical Inference Course Project - Part 1: Simulation Exercise Instructions'
author: "Omer Shechter"
date: "October 13, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This project you will investigate the exponential distribution in R and compare it
with the Central Limit Theorem. The exponential distribution can be simulated in
R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential 
distribution is 1/lambda and the standard deviation is also 1/lambda. 
Set lambda = 0.2 for all of the simulations. You will investigate the 
distribution of averages of 40 exponentials.




## Simulations
This part includes the simulation itself. 
Create a simulation vector of 1000 simulation, 
and calculating the mean of 40 samples in each simulation. 

Load required  libraries 
```{r}
library(ggplot2)
library(gridExtra)
```

Create the actual simulation :
Using the apply function to create 1000 simulation of 40 sampling and calculating the mean.
```{r}
set.seed(56)
nosim<-1000
n<-40
lambada<-0.2
sim_vect = c(apply(matrix(rexp(nosim * n,lambada), nosim), 1, mean))
```



## Sample Mean versus Theoretical Mean
Distribution for the mean of random samples
The expected value of the mean of the distribution of means = expected value of the sample mean =
population mean
* E[¯X ] = µ

In this case Expected mean = µ = 1/lambada  : 5 

Calculate the simulated sample mean:
```{r}
exp_mean<-mean(sim_vect)
print(exp_mean)
```


plot the distribution of sample means as n increases, and compare it to the population mean.
```{r fig.width=11,fig.height=11}
#calculate sample  means for different size of n
means <- cumsum(sim_vect) / (1 : nosim)
#plot sample size vs. sample mean as a function of the number of samples 
theme_update(plot.title = element_text(hjust = 0.5))
#plot using ggplot
g <- ggplot(data.frame(x = 1 : nosim, y = means), aes(x = x, y = y))
g <- g + geom_hline(yintercept = 0) + geom_line(size = 2)
g <- g + labs(x = "Number of observiations", y = "Cumulative mean")
g <- g + ggtitle(" Exponential distribution")
g <- g + geom_hline(yintercept=5, linetype="dashed", color = "red",size=1)
g <- g + ylim(3,7)
#g <- g + scale_y_continuous(breaks=c(3,4,5,6,7))
g <- g +  theme(axis.text = element_text(size = 14), 
        axis.title = element_text(size = 18),
        plot.title = element_text(size = 25, face = "bold"))
g

```
Summary: 
As it can be seen from the results above 
The sample mean provides an accurate estimator to the general population mean. 


## Sample Variance versus Theoretical Variance
Variance of distribution of ¯X = Sigma^2/n
estimate of variance = S2/n

In our example the Variance estimator is 
(1/Lambada/sqrt(n))^2 = 25/40 = .625

Calculate the simulated variance
```{r}
e_var<-(1/lambada/sqrt(n))^2
var(sim_vect)
```

Plot the sample variance as n increases and compare it to the population vars
```{r fig.width=11,fig.height=11}
#calculate sample  vars for different size of n
e_mean<-mean(sim_vect)
cvars <- cumsum((sim_vect-e_mean)^2)/seq_along((sim_vect-1))
#plot sample size vs. sample's var
theme_update(plot.title = element_text(hjust = 0.5))

g <- ggplot(data.frame(x = 1 : nosim, y = cvars), aes(x = x, y = y))
g <- g + geom_hline(yintercept = 0) + geom_line(size = 2)
g <- g + labs(x = "Number of observiations", y = "Cumulative vars")
g <- g + ggtitle(" Exponential distribution")
g <- g + geom_hline(yintercept=e_var, linetype="dashed", color = "red",size=1)
g <- g + scale_y_continuous(breaks=seq(0, 1, length.out = 11))
g <- g +  theme(axis.text = element_text(size = 14), 
        axis.title = element_text(size = 18),
        plot.title = element_text(size = 25, face = "bold"))
g

```


## Distribution
Plot a histogram of the density of the sample's mean 
Show that it is close to the normal distribution. 
```{r fig.width=11,fig.height=11}
g <- ggplot(data.frame(mu = sim_vect),aes(x=mu))
g <- g + geom_histogram(aes(y=..density..),color="black",fill="lightgreen",binwidth = .05)
g <- g + stat_function(fun=dnorm , color="black",args=list(mean=5,sd=sqrt(1/lambada/sqrt(n))^2))
g <- g + labs(x = "Means", y = "Density")
g <- g + ggtitle(" Density Histogram of multiple \n 40 Samples simulation  of the Exponential distribution")
g <- g + geom_vline(xintercept=1/lambada, linetype="dashed", color = "red",size=1)
g <- g +  theme(axis.text = element_text(size = 14), 
        axis.title = element_text(size = 16),
        plot.title = element_text(size = 16, face = "bold"))

g

```
Conclusion:
The above three sections give a good indication that the mean of samples (using R simulation) of a population that is IID with exponential distribution behave like a normal distribution.
The expected mean is 1/lambada, and the variance is  (1/Lambada/sqrt(n))^2.